---
title: 'Prophet package '
author: "Lucas"
date: "30 novembre 2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview and definition 

## What is Prophet ? 

Prophet is an open source package developed by Facebook Data Science Team in order 
to resolve business problems related to time series forecasting. 
Prophet is a procedure to forecast time series based on an additive model where 
non-linear trends are fit with yearly, weekly, and daily seasonality, plus
holiday effects. It works best with time series that have strong seasonal 
effects and several seasons of historical data. 
Prophet is robust to missing data and shifts in the trend, and typically 
handles outliers well.

Prophet is optimized for the business forecast tasks which typically 
have any of the following characteristics:

- hourly, daily, or weekly observations with at least a few months (preferably a year) of history
strong multiple “human-scale” seasonalities: day of week and time of year
- important holidays that occur at irregular intervals that are known in advance (e.g. the Super Bowl)
- a reasonable number of missing observations or large outliers
- historical trend changes, for instance due to product launches or logging changes
- trends that are non-linear growth curves, where a trend hits a natural limit or saturates

Prophet’s default settings can produce forecasts that are often accurate 
as those produced by skilled forecasters, with much less effort. With Prophet, 
you are not stuck with the results of a completely automatic 
procedure if the forecast is not satisfactory — an analyst with no training 
in time series methods can improve or tweak forecasts using 
a variety of easily-interpretable parameters. By combining automatic forecasting 
with analyst-in-the-loop forecasts for special cases, it is possible to 
cover a wide variety of business use-cases.

Therefore, the goal behind Prophet is to "make it easier for experts and
non-experts to make high-quality forecasts that keep up with demand."


## Real world Use caseof Prophet (some example)

- Prophet is used in many applications across Facebook for producing reliable forecasts for planning and goal setting
- forecast how many cars must be produce monthly to respond to the global demand 
- forecast how many PS5 will be sold over years to prevent over producing 

##How to install prophet 
first, you need to install Rtools
(follow the step on the CRAN) : https://cran.r-project.org/bin/windows/Rtools/

\newpage


# A simple model with prophet  

Before using Prophet you need to install it. 
It is a CRAN Package, so you can use install.packages(<package_name>)

```{r}
#install.packages('prophet')
```

Now that the package is installed, let's have a quick overview of how to use prophet in R. 
We will use a csv file available here : https://github.com/facebook/prophet/blob/master/examples/example_wp_log_peyton_manning.csv

This csv contains how many person saw Peyton Manning wikipedia page express in log. 

First step is to import the package and the csv file

The csv file must have specific columns names. 
This format must be respected otherwise prophet will not work. 

- ds reflect the date (YYYY-MM-DD, or YYYY-MM-DD HH:MM:SS) 
- y represent the output in this date (numerical values)

```{r}
library(prophet)
df <- read.csv("peyton_wikipedia_page.csv")
head(df)
```

We call the prophet function and we pass our dataframe as the first argument.
There are additional arguments that enable how prophet fits the data. 
We will see some of the in the next part. 

```{r}
m <- prophet(df)
```

The make_future_dataframe function takes the model object and a number of periods to forecast.  By default it will also include the historical dates so we can evaluate in-sample fit.

```{r}
future <- make_future_dataframe(m, periods = 365)
tail(future)
```

We use the predict() function to forecast on the date that we built using make_future_dataframe. 
It takes 2 arguments the "m" our model and "future" created from the previous function.
predict() return a dataframe with different information. 
here are some of them : 

- yhat wich is our forecast
- yhat_lower and yhat_upper which represent our uncertainty interval
  
```{r}
forecast <- predict(m, future)
head(forecast[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')])
```

You can plot the forecast, by passing in the model and the forecast dataframe.
You will have the 1st part of the plot that represent past data and the 2nd part
which is the forecast with the uncertainty interval.
```{r}
plot(m,forecast)
```

Finnaly,the prophet_plot_components() break down the forecast into 3 components.
The trend, the week seasonality, and the year seasonality.
.
```{r}
prophet_plot_components(m, forecast)
```

\newpage

# Intuition Behind Prophet 

As said earlier Prophet is a procedure for forecasting time series data based on
an additive model where non-linear trends are fit with yearly, weekly, and daily 
seasonality, plus holiday effects. 
But how does prophet work ? 
The prophet equation relies the deconstruction of a time series 
into several components, each representing one of the underlying categories of patterns. 
Prophet is based on 3 mains components : trend, seasonality and holidays.
In its simplest forms we can write the prophet equation as : 
y(t) = g(t) + s(t) + h(t) + e(t)
where : 
- g(t) trend  (i.e. growth over time)
- s(t) seasonality (i.e. weekly, monthly, yearly)
- h(t)ties in effects of holidays (on potentially irregular schedules >= 1 day(s))
- e(t)represents the errors

What prophet does is “framing the forecasting problem as a curve-fitting exercise”
rather than looking explicitly at the time based dependence of each observation.
 
# Some Maths for a better understanding 

Let's see in more details the Prophet equation. 
y(t) = g(t) + s(t) + h(t) + e(t)

We are going to see how the different parameter are modeling within Prophet. 

We will use the example_wp_log_R.csv dataset to illustrate 
each function and parameter of Prophet. 

example_wp_log_R.csv contains the informations of the number of people who have visited 
the wikipedia R page a specific day. 

## The trend g(t) 

Prophet allows 2 types of modeling for the trend. 
- piecewise linear model
- saturating growth model 


### Piecewise linear model

linear trend model is a simple Piecewise Linear Model with a constant rate of growth.
It is best suited for problems without a market cap or other max in sight.
Modeling the linear trend is easily realized with Prophet. 

g(t) = (k+a(t)T )t+(m+5T°()° 
where: 

where:
k is the growth rate
has the rate adjustments
m is the offset parameter
and, to make the function continuous,  is set to:

By default, Prophet is set to do its forecast using a linear model. 

We are first going to build the model using a linear model trend. 

```{r}
library(prophet)

df <- read.csv("log_wikipedia_R.csv")

#we specify growth = linear but you can also not specify (default setting)
m <- prophet(df, growth = "linear")

#we use the same process for the forecasting part 

future <- make_future_dataframe(m, periods = 1826)

forecast <- predict(m,future)

plot(m,forecast)

#we will store our plot 
linear <- plot(m,forecast)
```

### Saturating growth model 

If there is a saturation problem for example number of people with access to internet. 

g(t) equation is : g(t)=C/(1+exp(-k(t-m)))

Where : 
C = caring capacity 
k = growth rate 
m = offset parameter 

We can see from the previous plot that the forecast in terms of uncertainty is
not so good. 

When forecasting growth, there is usually some maximum achievable point: total 
market size, total population size, etc. This is called the carrying capacity, 
and the forecast should saturate at this point. 

let's modify our 1st model and precise the maximun capacity. the point at which 
the model should saturate.

the carrying capacity does not have to be constant if me size of the market is 
increasing the the capacity can be increasing to. 

for example have a restaurant with an area of 50m2 which can welcome 30 peoples.
You buy a local to expand your restaurant. Now the welcoming area is 70m² which means 
that you can welcome more people. 


```{r}
library(prophet)

df <- read.csv("log_wikipedia_R.csv")

#We must specify the carrying capacity in every row in the dataframe
#the carrying capacity is usually set using data expertise about the market size
df$cap <- 8.5 

#now we specify growth to logistic 
m <- prophet(df, growth = "logistic")

#we use the same process for the forecasting part 

future <- make_future_dataframe(m, periods = 1826)

#the capacity here is set to 8.5
future$cap <- 8.5 

forecast <- predict(m,future)

plot(m,forecast)

#we will store our plot 
growth <- plot(m,forecast)
```

## Seasonality s(t)

The seasonal component s(t) provides a adaptability to the model
by allowing periodic changes based on sub-daily, daily, 
weekly and yearly seasonality.

Prophet detects changepoints and allow the trend to adapt appropriately.
The locations of the signification changepoints can be visualized with : 

```{r}
plot(m,forecast) + add_changepoints_to_plot(m)
```
You can also adjust the strength of the sparse prior using the input argument 
changepoint_prior_scale. By default, this parameter is set to 0.05. Increasing it will make the trend more flexible:
(Decreasing it will make the trend less flexible:) 

```{r}
m <- prophet(df, changepoint.prior.scale = 0.5)
forecast <- predict(m, future)
plot(m, forecast)
```


## holydays and events h(t)

If there are specific event you want to model there are multiple choice.

we will see 2 of them : 

create a dataframe that includes the date you want to include. 
Use the built-in holidays country-specific collection

 
```{r}
#library(dplyr)

#playoffs <- data_frame(holiday = 'playoff',ds = as.Date(c('2008-01-13', '2009-01-03', #'2010-01-16','2010-01-24', '2010-02-07', '2011-01-08','2013-01-12', '2014-01-12', '2014-01-19',
#                 '2014-02-02', '2015-01-11', '2016-01-17',
#                 '2016-01-24', '2016-02-07')),
#  lower_window = 0,
#  upper_window = 1
#)

#holidays <- bind_rows(playoffs)

#library(prophet)
#df <- read.csv("peyton_wikipedia_page.csv")




#including the holidays we create previously
#m <- prophet(df,holidays =  holidays)

#m <- add_country_holidays(m, country_name =  'US')

#m <- fit.prophet(m, df)

#you can see the holiday included 
#m$train.holiday.names



#we use the same process for the forecasting part 

#forecast <- predict(m, future)
#prophet_plot_components(m, forecast)

```


## time series forecastiong with prophet 


\newpage


\newpage

# hyperparameter and how to tune the model (specifity in different work environement)
# Prophet pros and cons 

# Bibliography - usefull link - and data source 
https://facebook.github.io/prophet/docs/quick_start.html#r-api
https://facebook.github.io/prophet/
https://medium.com/future-vision/intro-to-prophet-9d5b1cbd674e
https://ledatascientist.com/facebook-prophet-la-prevision-a-grande-echelle/
https://towardsdatascience.com/a-quick-start-of-time-series-forecasting-with-a-practical-example-using-fb-prophet-31c4447a2274
https://mode.com/example-gallery/forecasting_prophet_r_cookbook/
https://peerj.com/preprints/3190/
https://research.fb.com/blog/2017/02/prophet-forecasting-at-scale/

